Linear regression for approximating real and synthetic data

### 2022-10-07 doc/paper.tex
More paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
{\it We use Ordinary Least Squares, Ridge and Lasso regression on a synthetic dataset generated by the Franke Function, and on a real dataset of elevation data from a terrain in Norway. We explore different regularization values and polynomial degrees and find that that MODEL performs best, with a MSE of NUMBER for polynomial degree 6. We find that linear regression performs well on the synthetic data, but falls short when approximating elevation data with large variances in height.
}
\end{abstract}

@@

@@

@@

Looking at Lasso regression, we find that it has the same capability to reduce overfitting as Ridge, as can be seen in Fig.~\ref{frankelassodifflambdas}.

\begin{table}[t]
\caption{Best MSE for Ridge regression and OLS with 400 datapoints}
\begin{center}
\label{ols_vs_ridge_table}
\begin{tabular}{c | l l l}
% & & \\ % put some space after the caption
% \hline
Regression & Std of noise & Best MSE & Best degree \\
\hline
OLS & 0.05 & 0.0038302 & 7 \\
Ridge & 0.05 & 0.0037999 & 7 \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

Turning our attention to Lasso regression, we also find that overfitting can be avoided with a good choice of $\lambda$. Comparing the three methods, we see that for noisy target data, Ridge and Lasso are superior to OLS, as seen in Fig. NUMBER.

@@

\begin{table*}
\caption{This table shows every applicable parameter for every figure, so that they may be easily reproduced. Every plot uses the random seed 42069}
\begin{center}
\label{allparamstable}
\begin{tabular}{c | l l l l l l l}
% & & \\ % put some space after the caption
% \hline
Figure & Highest $N$ & Std of noise & Number of datapoints & Bootstraps & K-folds & $\lambda$ & Scaling \\
\hline
\ref{frankenonoise} & 10 & 0 & 400 & - & - & - & None\\
\ref{frankenonoisemse} & 25 & 0 & 400 & 100 & - & - & None\\
\ref{frankenoisemse} & 25 & 0.05 & 400 & 100 & - & - & None\\
\ref{frankiebiasvariance} & 25 & 0.05 & 100 & 100 & - & - & None\\
\ref{frankiedifflambdas} & 25 & 0.05 & 100 & 100 & - & variable & Centering\\
\ref{frankienoisecomparison} & 25 & variable & 400 & 100 & - & - & None\\
\ref{frankiehigherdatapoints} & 30 & 0 & 10000 & 100 & - & - & None\\
\hline
\end{tabular}
\end{center}
\end{table*}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
{\it We use Ordinary Least Squares, Ridge and Lasso regression on a synthetic dataset generated by the Franke Function, and on a real dataset of elevation data from a terrain in Norway. We explore different regularization values and polynomial degrees and find that that all methods perform similarly on both sets of data. We find that linear regression performs well on the synthetic data, but falls short when approximating elevation data with large variances in height.
}
\end{abstract}

@@

% \begin{figure} 
% \centerline{\includegraphics[width=3.25in]{figure/frankenoisybeta.png}}
% \caption{Bias, variance and test MSE plotted over polynomial degree for our OLS prediction on the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$}
% \label{frankenoisybeta}
% \end{figure}

@@

\begin{table}[t]
\caption{Best MSE for Ridge regression and OLS with 400 datapoints}
\begin{center}
\label{ols_vs_ridge_table}
\begin{tabular}{c | l l l l}
% & & \\ % put some space after the caption
% \hline
Regression & Std of noise & MSE & Degree & $\lambda$ \\
\hline
OLS & 0.05 & 0.0038302 & 7 & - \\
Ridge & 0.05 & 0.0037999 & 7 & $10^{-8}$ \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

Looking at Lasso regression, we find that it has the same capability to reduce overfitting as Ridge, as can be seen in Fig.~\ref{frankelassodifflambdas}. However, looking at Table.~\ref{ols_vs_ridge_vs_lasso_table}, we see that Lasso falls short when finding a good MSE. Intuitively, it does not make sense for Lasso to be worse than OLS, because Lasso with $\lambda = 0$ is equivalent to OLS. But we cannot avoid considering the fact that Lasso does not have an analytical solution, and as such we must use numerical methods to find the optimal parameters. Using scikit's implementation, we have not been able to get the numerical method to converge, which means that we have not found a global minimum.\footnote{Scikit explicitly mentions that using $\lambda = 0$ does not converge well with their implementation} For this reason, Lasso regression has not given satisfying results.

\begin{table}[t]
\caption{Best MSE for OLS, Ridge and Lasso with 400 datapoints}
\begin{center}
\label{ols_vs_ridge_vs_lasso_table}
\begin{tabular}{c | l l l l}
% & & \\ % put some space after the caption
% \hline
Regression & Std of noise & MSE & Degree & $\lambda$ \\
\hline
OLS & 0.05 & 0.0038302 & 7 & - \\
Ridge & 0.05 & 0.0037999 & 7 & $10^{-8}$ \\
Lasso & 0.05 & 0.0095846 & 10 & $2.5118 \cdot 10^{-6}$ \\
\hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{figure} 
\centerline{\includegraphics[width=3.25in]{figure/frankeridgebetas.png}}
\caption{Beta progression for different values of lambda in Ridge regression. Note the different scales on each of the plots, showing how increasing the regularization parameter shrinks the betas}
\label{frankeridgebetas}
\end{figure}

@@

\begin{table*}
\caption{This table shows every applicable parameter for every figure, so that they may be easily reproduced. Every plot uses the random seed 42069}
\begin{center}
\label{allparamstable}
\begin{tabular}{c | l l l l l l l}
% & & \\ % put some space after the caption
% \hline
Figure & Highest $N$ & Std of noise & Number of datapoints & Bootstraps & K-folds & $\lambda$ & Scaling \\
\hline
\ref{frankenonoise} & 10 & 0 & 400 & - & - & - & None\\
\ref{frankenonoisemse} & 25 & 0 & 400 & 100 & - & - & None\\
\ref{frankenoisemse} & 25 & 0.05 & 400 & 100 & - & - & None\\
\ref{frankiebiasvariance} & 25 & 0.05 & 100 & 100 & - & - & None\\
\ref{frankiedifflambdas} & 25 & 0.05 & 100 & 100 & - & variable & Centering\\
\ref{frankeridgebetas} & 25 & 0.05 & 100 & - & - & variable & Centering\\
\ref{frankenoisybeta} & 25 & 0.05 & 400 & - & - & - & Centering\\
\ref{frankienoisecomparison} & 25 & variable & 400 & 100 & - & - & None\\
\ref{frankiehigherdatapoints} & 30 & 0 & 10000 & 100 & - & - & None\\
\hline
\end{tabular}
\end{center}
\end{table*}

@@

\begin{figure*}
% \begin{figure}[t]
\centerline{\includegraphics[width=3.25in]{figure/frankenoisybeta.png}}
\caption{Beta progression for OLS}
\label{frankenoisybeta}
\end{figure*}
